Appendix 4: example of index standardization of fishery-independent survey data with inlabru

### 2021-12-13 doc/appendix-pcod-inlabru.Rmd
inlabru not working with tweedie

This model will attempt to replicate the model fit in Appendix 1 as an example of how index standardization can also be achieved with inlabru.
We will start by building an identical mesh using INLA.
These same functions that are used internally in the sdmTMB `make_mesh()` function.

@@

```{r fit-pcod, warning=FALSE, message=FALSE}
m <- lm(present ~ -1 + as.factor(year), data = pcod)
model_mat <- model.matrix(m)
colnames(model_mat) <- paste0("x", unique(pcod$year))
dat <- cbind(pcod, model_mat)
```

@@

```{r}
components <- present ~ -1 + x2003 + x2004 + x2005 + x2007 +
  x2009 + x2011 + x2013 + x2015 + x2017 +
  spatrf(main = coordinates, model = spde)
# spatrf(main = coordinates,model = spde, group=year, ngroup=ncol(model_mat),
#       control.group = list(model = "iid"))

fit <- bru(
  components = components,
  family = "binomial", data = dat
)
bru_coefs <- fit$summary.fixed[, c("mean", "sd")]
```

@@

sdmtmb_fit <- sdmTMB(present ~ -1 + as.factor(year),
  data = pcod,
  time = "year", mesh = mesh,
  family = binomial(link = "logit"),
  spatiotemporal = "off",
  priors = sdmTMBpriors(
    matern_st = pc_matern(
      range_gt = 10, range_prob = 0.05,
      sigma_lt = 2, sigma_prob = 0.05
    )
  )
)
sdmtmb_coefs <- tidy(sdmtmb_fit, effects = "fixed")
```

@@

```{r}
components <- present ~ 0 +
  fac(main = year, model = "factor_full") +
  spatrf(main = coordinates, model = spde)

fit2 <- bru(
  components = components,
  family = "binomial", data = dat
)
bru_coefs2 <- fit2$summary.random$fac[, c("mean", "sd")]
```

@@

```{r}
plot(bru_coefs$mean, bru_coefs2$mean)
```

>>>>

This model will attempt to replicate the model fit in Appendix 1 as an example of how index standardization can also be achieved with inlabru. 
Once again we will use a data set built into the package: trawl survey data for Pacific Cod in Queen Charlotte Sound.
The density units are kg/km^2^. 
This can be converted to tonnes per km2 by dividing by 1000, which can sometimes scale better.

```{r}
pcod$density_tpkm <- pcod$density/1000
```

We will start by building an identical mesh using INLA.
Here, X and Y are coordinates in UTM zone 9.
These same functions that are used internally in the sdmTMB `make_mesh()` function.

@@

```{r fit-pcod, warning=FALSE, message=FALSE}
m <- lm(density_tpkm ~ -1 + as.factor(year), data = pcod)
model_mat <- model.matrix(m)
colnames(model_mat) <- paste0("x", unique(pcod$year))
dat <- cbind(pcod, model_mat)
```

@@

```{r}
components <- density_tpkm ~ -1 + x2003 + x2004 + x2005 + x2007 +
  x2009 + x2011 + x2013 + x2015 + x2017 +
  # spatrf(main = coordinates, model = spde)
  spatrf(main = coordinates, model = spde, group=year, ngroup=ncol(model_mat),
      control.group = list(model = "iid"))

fit <- bru(
  components = components,
  family = "tweedie",
  # family = "binomial",
  options = bru_options(bru_verbose = TRUE), 
  data = dat
)
bru_coefs <- fit$summary.fixed[, c("mean", "sd")]
```

@@

sdmtmb_fit <- sdmTMB(density_tpkm ~ -1 + as.factor(year),
  data = pcod,
  time = "year", mesh = mesh,
  family = tweedie(link = "log"),
  # family = binomial(link = "logit"),
  # priors = sdmTMBpriors(
  #   matern_st = pc_matern(
  #     range_gt = 10, range_prob = 0.05,
  #     sigma_lt = 2, sigma_prob = 0.05
  #   )
  # ),
  spatiotemporal = "IID"
)
sdmtmb_coefs <- tidy(sdmtmb_fit, effects = "fixed")
```

@@

```{r}
# components <- present ~ 0 +
#   fac(main = year, model = "factor_full") +
#   spatrf(main = coordinates, model = spde)
# 
# fit2 <- bru(
#   components = components,
#   family = "binomial", data = dat
# )
# bru_coefs2 <- fit2$summary.random$fac[, c("mean", "sd")]
```

@@

```{r}
# plot(bru_coefs$mean, bru_coefs2$mean)
```
